\section{Conclusion and Future Work}\label{sec:conclusion}
I have presented my work on implementing and formalizing shortcut fusion of (Co)Church encodings as described by \cite{Harper2011}.
I have replicated Harper's work of Church and Cochurch encoded functions operating on leaf trees: \tt{between}, \tt{map}, \tt{filter}, and \tt{sum};  and shown the generalizability of his example by also implementing the functions on lists.
In doing so I discovered that in Haskell full fusion is not currently possible for the Cochurch encoded filter function.
GHC needs either proper loopification using join points\footnote{See \url{https://gitlab.haskell.org/ghc/ghc/-/issues/22227\#note_551000}}, or additional encoding techniques such as those described by \cite{Coutts2007}.

I benchmarked the performance of multiple different variants of the same pipeline: unencoded, hand-fused, Church fused, Cochurch fused, and GHC.List fused; where the (Co)Church fused pipelines had four variants: tail recursive, stream fused, neither, and both.
I discovered that changing the underlying datatype for Church encodings from List to Stream datatypes gave no performance improvement, for both tail and non-tail recursive implementations. Implementing tail recursion however did offer a speedup, for Cochurch encodings.
It was also faster to implement tail recrusion in addition to modifying the underlying type from List to Stream.
This was likely due to the improper loopification of the recursive coalgebra \tt{go}.
The fully fused (fastest) pipelines of both Church and Cochurch encodings were about as fast as the hand-fused and GHC.List fused pipelines; for some inputs the (Co)Church fusion was faster, for others the hand-fused/GHC.List fused.

I implemented Harper's description of Church and Cochurch encodings using Agda's dependent type system, using containers to represent strictly positive functors.
Before formalizing the proof of the shortcut fusion property, I first formalized all of the needed underlying category theory: the universal property of folds (i.e., initiality of initial algebras), the computation law, the reflection law, and the fusion property.
Using these, I formalized Harper's proofs of the Church and Cochurch encodings being faithful, showing that they are isomorphic to the datatype that they are encoding.
This came with one major caveat: The reliance on the free theorems of parametric functions, which was not provable in Agda.
There is recent work on this \textit{internalized parametricity} by \cite{Muylder2024}, which would make the free theorems provable from within Agda, dubbed Agda --bridges.
Finally, I implemented (Co)Church encoded versions of the following four functions \tt{between}, \tt{map}, \tt{filter}, and \tt{sum} and showed that their composition as (Co)Church encodings was equal to the hand fused function also presented.

These findings highlight the effectiveness and correctness of shortcut fusion techniques and show the promise of shortcut fusion: Reduce the computational overhead associated with functional programming while retaining its nice, compositional properties.

\subsection*{Future Work}
There are many future avenues that could be taken to continue my research:
\begin{itemize}[noitemsep]
    \item Tool (Language) improvements:
    \begin{itemize}[nolistsep]
        \item See if it is possible to implement warm fusion in Haskell or some other language as described by \cite{Launchbury1995}.
        \item Strengthen Agda's typechecker with respect to implicit parameters. Currently two variants of functional extensionality had to be defined to work around this.
        \item Investigate if creating a new programming language that has this fusion as a first-class feature can enable fusion to be compiled more efficiently and consistently.
    \end{itemize}
    \item Extensions of my work
    \begin{itemize}[nolistsep]
        \item Implement a bisimilarity relation for the coinductive \tt{M}/$\nu$  type in Agda to prove its terminality. After which modifying all the code resting on top of this proof to properly use this new relation.
        \item Investige whether it is to generalize the work of \cite{Coutts2007} to more datastrcutures, with a motivating example being Leaf Trees.
        \item Use Agda --bridges to see if it is possible to prove the free theorems currently postulated in my work.
    \end{itemize}
    \item Applications of my work:
    \begin{itemize}[nolistsep]
        \item Implement (Co)Church fused versions of Haskell's library functions.
        \item Merge into Agda the Church and Cochurch encodings, as well as the bisimilarity across the guarged \tt{M} type.
    \end{itemize}
\end{itemize}
